---
title: "Chocolate Chip Cookies"
execute:
  error: true
author: "Your Name"
output: html_document
---

## Reading In the Data

First, read in the CSV data of cookie ingredients.
Make sure that your end-result data has appropriate types for each column - these should match the types provided in the documentation in the README.md file.

```{r}
cookie_data <- read.csv("~/stat251-week4/choc_chip_cookie_ingredients.csv")
```

```{python}
import pandas as pd
cookie_data = pd.read_csv("~/stat251-week4/choc_chip_cookie_ingredients.csv")
print(cookie_data)
```


## Exploratory Data Analysis

Exploratory data analysis is the process of getting familiar with your dataset. To get started, [this blog post](https://www.mrdbourke.com/a-gentle-introduction-to-exploratory-data-analysis/) provides a nice checklist to get you thinking:

> 1.  What question(s) are you trying to solve (or prove wrong)?
> 2.  What kind of data do you have and how do you treat different types?
> 3.  What's missing from the data and how do you deal with it?
> 4.  Where are the outliers and why should you care about them?
> 5.  How can you add, change or remove features to get more out of your data?

### Generating Questions

Generate at least 5 questions you might explore using this database of cookie ingredients.

1. What is the average amount of each ingredient needed?
2. What's missing from the cookie data?
3. Are there any outliers in the amount of ingredients used?
4. Will the different units of milk used affect how we interpret certain data?
5. Is it necessary to keep any ingredients that only show up once or a few times in the dataset?                                                                      


### Skimming the Data

One thing we often want to do during EDA is to examine the quality of the data - are there missing values? What quirks might exist in the dataset?

The `skimr` package in R, and the similar `skimpy` package in python (which has a much better name, in my opinion), can help provide visual summaries of the data. 

Install both packages, and read the package documentation ([R](https://cran.r-project.org/web/packages/skimr/vignettes/skimr.html), [Python](https://pypi.org/project/skimpy/)).

[Part 1] Use each package and generate summaries of your data that require the use of at least some non-default options in each package's `skim` function.


```{r}
library("skimr")

# Only Ingredient, Rating, Quantity, and Unit columns
skim(cookie_data, Ingredient, Rating, Quantity, Unit)


```

```{python}
from skimpy import skim
import pandas as pd

# only numeric columns
skim(data.select_dtypes(include='number'))

```

[Part 2] Write 1-2 sentences about what you can tell from each summary display you generate. Did you discover anything new about the data?

There are 1010 missing values in the Rating column, there are 1990 rows, 2 numeric columns use floats and 1 numeric column uses integers.


### Generating Tables

Another useful technique for exploratory data analysis is to generate summary tables. 
You may want to use the `dplyr` package in R (`group_by` or `count` functions), as well as the `groupby` and `count` methods in Pandas. [Python example](https://sparkbyexamples.com/pandas/pandas-groupby-count-examples/), [R example](https://dplyr.tidyverse.org/reference/count.html)

[Part 1] Using R and Python, generate a table that shows what **proportion** of recipes contain each type of ingredient, for the most common 20 ingredients.
```{r}
library("dplyr")

ingredient_proportions <- cookie_data %>%
  distinct(Recipe_Index, Ingredient) %>% 
  count(Ingredient, name = "recipe_count") %>%
  mutate(proportion = recipe_count / n_distinct(cookie_data$Recipe_Index)) %>% # Proportion
  arrange(desc(recipe_count)) %>% # In descending order 
  slice_head(n = 20) # Top 20 ingredients

print(ingredient_proportions)

```

```{python}
import pandas as pd

# Removes duplicate ingredients in the same recipe
distinct_ingredients = cookie_data.drop_duplicates(subset=["Recipe_Index", "Ingredient"])

# Count of how many recipes each ingredient appears in
counts = (
  distinct_ingredients
  .groupby("Ingredient")["Recipe_Index"]
  .nunique()
)

ingredient_prop = (
  counts
  .to_frame(name = "recipe_count")
  .assign(
    proportion=lambda x: x["recipe_count"] / distinct_ingredients["Recipe_Index"].nunique()
  )
  .sort_values("recipe_count", ascending=False)
  .head(20)
)

print(ingredient_prop)
```
[Part 2] Print out a character string that lists all of the ingredients that do not appear in at least 20 recipes.

```{r}
library(dplyr)

# Count of recipes each ingredient is in
ingredient_counts <- cookie_data %>%
  distinct(Recipe_Index, Ingredient) %>% 
  count(Ingredient, name = "recipe_count")

# Filter ingredients that do not appear in at least 20 recipes
rare_ingredients <- ingredient_counts %>%
  filter(recipe_count < 20) %>%
  pull(Ingredient)

# Convert to a single character string
rare_ingredients_string <- paste(rare_ingredients, collapse = ", ")

rare_ingredients_string

```

### Visualization

Using whatever plotting system you are comfortable with in R or python, see if you can create a couple of useful exploratory data visualizations which address one of the questions you wrote above - or another question which you've come up with as you've worked on this assignment.

[Part 1] Create at least one plot (it doesn't have to be pretty) that showcases an interesting facet of the data.

```{python}
import matplotlib.pyplot as plt

egg_data = cookie_data[cookie_data["Ingredient"] == "egg"]

plt.figure(figsize=(8, 6))
plt.scatter(egg_data["Recipe_Index"], egg_data["Quantity"],
            color="darkblue", alpha=0.7)

plt.title("Quantities of Egg Used Across Recipes")
plt.xlabel("Recipe Index")
plt.ylabel("Egg Quantity")

plt.tight_layout()
plt.show()


```


[Part 2] Write 2-3 sentences about what you can learn from that plot and what directions you might want to investigate from here.

I learned from this plot that there is one outlier that uses 16 eggs, every other recipe stays within 1-8 eggs. Using this I may also want to know other proportions like the amount of flour or baking powder (any dry ingredient) used across all recipes that use them as well.